---
output:
   word_document:
      reference_docx: working_docx_template_nonumbers.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      # - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      # - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
header-includes:
- \usepackage{amsmath}
- \numberwithin{equation}
- \usepackage{indentfirst}
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

```{r import general objects, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
here::i_am("doc/Junker_temp-energy-flux_appendixS3.Rmd")
source("./packages.R")

```

Appendix S3: Supporting information for Junker, J. R., W. F. Cross, J. M. Hood, J. P. Benstead, A. D. Huryn, D. Nelson, J. S. Ólafsson, and G. M. Gíslason, "*Environmental warming increases the importance of high-turnover energy channels in stream food webs*" for review and publication in *Ecology*.

## Appendix S3: Section S1: Trophic basis of production workflow and assumptions 

We estimated organic matter consumption (g m^-2^ y^-1^) for all taxa using the trophic basis of production method [@benke1980; @benke1997]. This method combines empirical estimates of secondary production and diet composition with assumed diet-specific assimilation, *AE*, and net production efficiencies, *NPE*, to calculate consumption of each diet item required to support secondary production of a consumer.

To propagate uncertainty in consumption estimates, we combined bootstrapped secondary production estimates calculated from resampling replicated surber samples across sampling dates with diet composition estimates from a hierarchical Bayesian model and assumed distributions of ecological efficiences. This supplement is a guide to the general workflow we used to estimate uncertainty in consumption estimates in consumers in this study. We walk through a single example here to 

\newpage

```{r TBP workflow}
knitr::include_graphics(here::here("doc/revision/TBP-supplemental_flat.tif"), rel_path = FALSE)
```

# References